\chapter{Installing Dynare} \label{ch:inst}

\section{Dynare versions}
Three versions of Dynare exist: one for \textbf{Matlab}, one for \textbf{Scilab} and one for \textbf{Gauss}. The first benefits from ongoing development and is the most popular. Development of the Scilab version stopped after Dynare version 3.02 and that for Gauss after Dynare version 1.2. \\

This User Guide will exclusively \textbf{focus on the Matlab version of Dynare}. For the installation procedure for the Scilab or Gauss versions of the program, please see the \href{http://www.cepremap.cnrs.fr/juillard/mambo/index.php?option=com_content&task=view&id=51&Itemid=84}{Reference Manual}. Note, though, that the Dynare syntax remains mostly unchanged across the Matlab, Scilab or Gauss versions, for those features common to the three versions. \\

You may also be interested by another version of Dynare, developed in parallel: \textbf{Dynare++}. This is a standalone C++ version of Dynare specialized in computing k-order approximations of dynamic stochastic general equilibrium models. See the \href{http://www.cepremap.cnrs.fr/juillard/mambo/index.php?option=com_content&task=view&id=53&Itemid=86}{Dynare++ webpage} for more information. 

\section{System requirements}
Dynare can run on Windows 98, 2000, NT and XP; at the time of writing, no experience was available on Vista (** check?). Dynare can also run on Unix-like operating systems, such as any Linux distribution, Solaris and, of course, Mac OS X. If you have questions about the support of a particular platform, feel free to write directly to Michel Juillard (michel.juillard"AT"ens.fr) or visit the \href{http://www.cepremap.cnrs.fr/juillard/mambo/index.php?option=com_forum&Itemid=95&page=viewforum&f=1}{\textbf{Dynare forums}}. \\

To run Dynare, it is recommended to allocate at least 256MB of RAM to the platform running Dynare, although 512MB is preferred. Depending on the type of computations required, like the very processor intensive Metropolis Hastings algorithm, you may need up to 1GB of RAM to obtain acceptable computational times. \\

\section{Installing Dynare}
\subsection{Installing on Windows}
The following assumes you have Matlab version 6.5.1 or later installed on your Windows system.\footnote{As of writing this Guide, Dynare is being developed on Matlab version 7. Nonetheless, great care is taken not to introduce features that would not work with reasonably recent versions of Matlab. However, Dynare requires at least the Matlab feature set of version 6.5.1, released September 22, 2003.}
\begin{enumerate}
\item Download the latest stable version of Dynare for Matlab (Windows) from the \href{http://www.cepremap.cnrs.fr/juillard/mambo/index.php?option=com_frontpage&Itemid=1}{Dynare website}. 
\item You will now have on your computer a .zip file which you should unzip. This will create a folder called, by default, Dynare and its version number, for example: Dynare\_v3.0 
\item This directory contains several sub-directories, among which (i) matlab, (ii) doc and (iii) examples. 
\item Place the Dynare folder (Dynare\_v3.0 in our example) in the c: directory and note that location. The easiest is probably to put it in the root of c: as in c:/dynare\_v3.0.
\item Start Matlab and use the menu File/Set-Path to add the path to the Dynare 
matlab subdirectory. Following our example, this would correspond to 
c:/dynare\_v3.0/matlab
\item Save these changes in Matlab and you're ready to go. (** doesn't work if dynare is put in ''program files''... bug?)
\end{enumerate} 

\subsection{Installing on UNIX}
** TBD - must recompile parser, need exact instructions. 
\subsection{Installing on Mac OSX}
** TBD - must recompile parser, need exact instructions. 

\section{Matlab particularities}

A question often comes up: what special Matlab toolboxes are necessary to run Dynare? In fact, no additional toolbox is necessary for running most of Dynare, except maybe for optimal simple rules (see chapter \ref{ch:ramsey}), but even then remedies exist (see the \href{http://www.cepremap.cnrs.fr/juillard/mambo/index.php?option=com_forum&Itemid=95}{Dynare forums} for discussions on this, or to ask your particular question). But if you do have the `optimization toolbox' installed, you will have additional options for solving for the steady state (solve\_algo option) and for searching for the posterior mode (mode\_compute option), both of which are defined later. 


